Verification of the CVM algorithm with a New Recursive Analysis Technique

### 2025-02-18 paper/sec/abstract.tex
slight edit

Central to our formalization is a new proof technique based on \emph{functional probabilistic invariants}, which allows us to derive concentration bounds using the Cram\'{e}r--Chernoff method.
This simplifies the formal analysis considerably compared to the original transformation-based proof.
Moreover, our technique opens up the possible algorithm design space; we demonstrate this by introducing and verifying with a new variant of the CVM algorithm that is total and unbiased---neither property is possessed by the original algorithm.
In this paper, we introduce the proof technique, describe its use in Isabelle/HOL to mechanize both versions of the CVM algorithm, and present a supporting formalized library on negative association used to verify the latter variant.
\end{abstract}

>>>>

Central to our formalization is a new proof technique based on \emph{functional probabilistic invariants}, which allows us to derive concentration bounds using the Cram\'{e}r--Chernoff method.
This simplifies the formal analysis considerably compared to the original transformation-based proof.
Moreover, our technique opens up the possible algorithm design space; we demonstrate this by introducing and verifying with a new variant of the CVM algorithm that is total and unbiased---neither property is possessed by the original algorithm.
In this paper, we introduce the proof technique, describe its use in Isabelle/HOL to mechanize both versions of the CVM algorithm, and present a supporting formalized library on negatively associated random variables used to verify the latter variant.
\end{abstract}
